\documentclass[11pt]{article}

% --- Packages ---
\usepackage[usenames, dvipsnames]{color} % Cool colors
\usepackage{enumerate, amsmath, amsthm, amssymb, mathrsfs, algorithm, algpseudocode, fontawesome, pifont, subfig, fullpage, csquotes, dashrule, tikz, bbm, booktabs, bm, hyperref}
\usepackage[framemethod=TikZ]{mdframed}
\usepackage[numbers]{natbib}
\usepackage[normalem]{ulem}

% --- Misc. ---
\hbadness=10000 % No "underfull hbox" messages.
\setlength{\parindent}{0pt} % Removes all indentation.

% -- Commands --
\input{commands}

\title{IACAP 2019 Notes \\ \Large{Mexico City, Mexico}}
\author{David Abel\footnote{\durl{http://david-abel.github.io}} \\ \durl{david_abel@brown.edu}}
\date{June 2019}

\begin{document}
\maketitle
\tableofcontents
\newpage


This document contains notes I took during the events I managed to make it to at (my first) conference of the International Association for Computing and Philosophy (IACAP), in Mexico City, Mexico. Please feel free to distribute it and shoot me an email at \durl{david_abel@brown.edu} if you find any typos or other items that need correcting. 


\section{Conference Highlights}

IACAP was a really fascinating experience--I came away with a mix of thoughts on the interplay between philosophy and computing, and the role computing is playing in our society at the moment (questions, I find, distinct from those typically raised at ML conferences). \\

One thing I loved about the conference was the scope of material: almost any topic/question was on the table to be explored. It's one thing that fundamentally separates CS from philosophy: philosophy is (often) concerned with asking questions in the right way and clarifying our inquiry, no matter the topic. ML/AI, at least at the moment, is usually focused on specific technical contributions (often a new algorithm, new theorems, experiments, and so on). Of course, these different artifacts serve different purposes, and both matter a great deal. To this end, the conference spanned a great deal of material, from non-classical logics, to self-driving cars, to meta-ethics, to the role computation plays (or should play) in the philosophy of science. \\

A few takeaways:
\begin{enumerate}
    \item Philosophy and computation (and, to me, specifically ML/AI) still have loads to share with one another.
    
    ($\ra$ But, doing so can be very challenging: different tools, methods, and objectives!)
    
    \item I really enjoyed John Weckert's talk ``Our Technology Fetish", discussing the world's current irrational obsession with development of technology (even in cases where we don't have an obvious use for such systems). In many ways it was a call for rethinking some of the values of what constitutes a flourishing society.
    
    \item Fiona McEvoy's talk also hit home: she discussed how automated systems are driving behavioral change through ``nudges" in online (and other settings)---it had flavors of Zeynep Tufecki's 2019 ICLR keynote. The message in both cases is for us to be wary of the kinds of psychological impact that subliminal interventions online can have on our communities, such as {\it fundamental} changes to our worldview and values.
    
    \item I *loved* how international and friendly the conference was: it's a tight knit community of folks that have been attending IACAP for awhile, so many have known each other for quite some time. It also drew in people from all over the world, so the perspectives shared about the role of technology were extremely diverse.
\end{enumerate}


% ------------
% -- Sunday --
% ------------
\newpage
\section{Wednesday June 5th}
\input{days/wednesday_6_5_19.tex}



% ------------
% -- Monday --
% ------------
\newpage
\section{Thursday June 6th}
\input{days/thursday_6_6_19.tex}



% ------------
% -- Tuesday --
% ------------
\newpage
\section{Friday June 7th}
\input{days/friday_6_7_19.tex}




% --- Bibliography ---
\newpage
\bibliographystyle{plainnat}
\bibliography{iacap}

\end{document}